
## IRT analysis

Item Response Theory (IRT) is a nice way to look at the informativeness of individual items with respect to a latent dimension. Again we infer here that there is a latent dimension that is being recovered (which may not always be true). But if there is one, we can use it to pick out bum items. 

### Sample (Crawling)

Sample example. Consider the "gateo" (crawling) category.

```{r}
gateo <- filter(d, category == "Gateo") %>%
  select(id, age, milestone, response) %>% 
  spread(milestone, response)

ltm::descript(select(gateo, -age, -id))
```

Fit simple Rasch model. 

```{r}
gateo_simple <- select(gateo, -age, -id)
fit1 <- rasch(gateo_simple, 
              constraint = cbind(length(gateo_simple) + 1, 1))
coef(fit1, prob = TRUE, order = TRUE)
plot(fit1)
```

Looks like the "backwards downstairs item" is very different. 

Check goodness of fit.

```{r}
GoF.rasch(fit1, B = 199)
```

Significance value suggests that the model doesn't fit that well. 

```{r}
margins(fit1)
```

Unconstrained Rasch. We need a discrimination parameter that's different than one. 

```{r}
fit2 <- rasch(gateo_simple)
coef(fit1, prob = TRUE, order = TRUE)
anova(fit1, fit2)
plot(fit2)
```

Now try an LTM.

```{r}
fit3 <- ltm(gateo_simple ~ z1)
anova(fit2, fit3)
coef(fit3, prob = TRUE, order = TRUE)
plot(fit3)
```

Item information curves. 

```{r}
plot(fit3, type = "IIC")
```

Adjust by dropping backwards item. 

```{r}
fit3 <- ltm(select(gateo_simple, -`Baja las escaleras gateando de reversa`, 
                   -`Cambia de dirección fácilmente al gatear`) ~ z1)
plot(fit3)
plot(fit3, type = "IIC")
```

### IRT Applied Broadly

Check on total test info across items. 

```{r}
info <- d %>%
  ungroup %>%
  select(id, milestone, category, response) %>%
  split(.$category) %>%
  map_df(function(x) {
    wide <- x %>%
      mutate(response = as.numeric(response)) %>%
      spread(milestone, response) %>%
      select(-id, -category) %>%
      data.frame 
    
    mod <- ltm(wide ~ z1)
    
    return(data_frame(category = x$category[1], 
                      info = information(mod, range = 1:ncol(wide))$InfoTotal))
  })


n_milestones <- cats %>%
  group_by(category) %>%
  summarise(n_milestones = n())

info <- left_join(info, n_milestones)

info$category <- fct_reorder(info$category, info$info)

ggplot(info, aes(x = category, y = info)) + 
  geom_point(aes(size = n_milestones)) + 
  coord_flip()
```

Now compute individual information for each question on each subtest. 

```{r}
m_info <- d %>%
  ungroup %>%
  select(id, milestone, category, response) %>%
  split(.$category) %>%
  map_df(function(x) {
    wide <- x %>%
      mutate(response = as.numeric(response)) %>%
      spread(milestone, response) %>%
      select(-id, -category) %>%
      data.frame 
    
    mod <- ltm(wide ~ z1)
    
    m_info <- data_frame(category = x$category[1], 
                         milestone = unique(x$milestone),
                         index = 1:length(unique(x$milestone)))
    
    m_info <- m_info %>%
      split(.$milestone) %>%
      map_df(function(y) {
        return(bind_cols(y, data.frame(info = 
                                         information(mod, 
                                                     range = 1:ncol(wide), 
                                                     items = y$index)$InfoTotal)))})
    
    return(m_info)
  })


info$category <- fct_reorder(info$category, info$info)

ggplot(info, aes(x = category, y = info)) + 
  geom_point(aes(size = n_milestones)) + 
  coord_flip()
```


